\section{Related Work}
\label{sec:Related}

ATOM \cite{srivastava1994atom} was one of the first and has remained one of the
more popular static binary instrumentation toolkits available. ATOM works in a
way that is conceptually similar to PEBIL; instrumentation is performed on the
compiled binary prior to runtime, meaning that any overhead due to code analysis
and code generation is incurred outside of the instrumented application's run
cycle. Unfortunately, ATOM is available only for the Alpha platform. Since this
processor is not being produced anymore, ATOM is no longer viable as a long-term
solution for those who wish to perform static, efficient instrumentation. 

Dyninst \cite{buck2000api} is a popular dynamic instrumentation toolkit that
uses a technique called code patching to perform instrumentation. Similar in
concept to what is done in PEBIL, this technique replaces an instruction from
the application with a jump instruction to a function call stub and
instrumentation code. The key difference between PEBIL and Dyninst is that
Dyninst performs all code patching at runtime instead of prior to runtime. This
has several advantages, including the ability to insert, remove and customize
instrumentation during runtime. But performing modification to the program at
runtime also may have a significant performance disadvantage, resulting in
inefficient execution of the instrumented application.

Pin \cite{luk2005pin} is another popular dynamic binary instrumentation toolkit
that uses a JIT-based (Just In Time compilation) approach to instrumentation.
This approach entails running the application on top of Pin, while Pin
intercepts the application at each natural control flow interruption in the
program to perform instrumentation on the next part of the program. For
efficiency, Pin performs many optimizations including caching these instrumented
sequences of code to allow for re-use, chaining instrumented sequences of code
together to avoid unnecessary tool intervention, and avoiding state protection
overheads whenever possible.

DynamoRIO \cite{bruening2004efficient} and Valgrind
\cite{nethercote2007valgrind} are two other dynamic binary instrumentation
toolkits that use a JIT-based approach to instrumentation and operate in a
similar fashion to Pin. These toolkits offer certain functionality that is not
available anywhere else. For example, Valgrind offers support for a feature
called shadow values \cite{nethercote2007shadow}, which can be used to create
instrumentation tools that are difficult to build without their
support. However, support of this kind entails a more heavyweight approach to
instrumentation that is unsatisfactory when efficiency is the primary goal.
